---
title: "Sequence Report"
author: "Julie Gilbert"
date: "6/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

# Report Overview

This report uses ```/SequenceSampleMetadata/FinalSummary/secret/full_compiled_data.csv``` as its data source. The goal of this report is to provide summary statistics on the COVID-19 sequencing status of Dr. Adam Lauring's research lab. 

This report was run on `r Sys.Date()`.

For questions, please contact Julie (Jules) Gilbert [juliegil@umich.edu].

--- 

```{r}
library(tidyverse)
library(reshape2)
library(lubridate)
library(gt)
```

```{r}

data_file <- read.csv("C:/Users/juliegil/Dropbox (University of Michigan)/MED-LauringLab/SequenceSampleMetadata/FinalSummary/secret/full_compiled_data.csv", colClasses = c("character"))

who_cw <- read.csv("C:/Users/juliegil/Documents/UofM_Work/SequenceCompilationCode/OutsidePipeline/who_variant_crosswalk.csv")

data_file$voi_voc_pangolin <- ifelse(grepl(who_cw$pango_lineage, data_file$pangolin_lineage), who_cw$pango_lineage, "Not VOI/VOC")

data_file$voi_voc_nextclade <- ifelse(grepl(who_cw$nextclade_clade, data_file$nextclade_clade), who_cw$nextclade_clade, "Not VOI/VOC")

data_file$voi_voc <- ifelse(data_file$voi_voc_pangolin == "Not VOI/VOC", "Not VOI/VOC", 
                            ifelse(data_file$voi_voc_nextclade == "Not VOI/VOC", "Not VOI/VOC", "VOI/VOC"))
#table(data_file$voi_voc)
```


## Complete Data Review

```{r}
### fix date formatting
data_file <- data_file %>% mutate(coll_date = case_when(grepl("/", coll_date) ~ as.POSIXct(coll_date, format = "%m/%d/%Y"), 
                                                        grepl("-", coll_date) ~ as.POSIXct(coll_date, format = "%Y-%m-%d"), 
                                                        T ~ NA_real_))


#data_file$coll_date <- as_date(data_file$coll_date)

data_file$PlateDate <- as_date(data_file$PlateDate)

data_file$counter <- 1
```


The full_compiled_data.csv dataset includes `r nrow(data_file)` sample records. This includes samples collected from `r min(data_file$coll_date, na.rm = TRUE)` to `r max(data_file$coll_date, na.rm = TRUE)`. The most recent samples were processed on `r max(data_file$PlateDate, na.rm = TRUE)`.

***Table 1.*** Sample Count from Received Sources

```{r}
sample_count <- data_file %>% group_by(received_source) %>% summarize(unique_samples = length(unique(sample_id)), total_samples = length(sample_id), unique_subjects = length(unique(subject_id)), min_collection = min(coll_date, na.rm = TRUE), max_collection = max(coll_date, na.rm = TRUE), min_plate_run = min(PlateDate, na.rm = TRUE), max_plate_run = max(PlateDate, na.rm = TRUE))

colnames(sample_count) <- c("Source", "Count of Unique Samples", "Count of Total Samples", "Count of Unique Subjects", "Min. Collection Date", "Max. Collection Date", "Min. Plate Run Date", "Max. Plate Run Date")

sample_count2 <- data_file %>% group_by(counter) %>% summarize(unique_samples = length(unique(sample_id)), total_samples = length(sample_id), unique_subjects = length(unique(subject_id)), min_collection = min(coll_date, na.rm = TRUE), max_collection = max(coll_date, na.rm = TRUE), min_plate_run = min(PlateDate, na.rm = TRUE), max_plate_run = max(PlateDate, na.rm = TRUE))

colnames(sample_count2) <- c("Source", "Count of Unique Samples", "Count of Total Samples", "Count of Unique Subjects", "Min. Collection Date", "Max. Collection Date", "Min. Plate Run Date", "Max. Plate Run Date")

sample_count2$Source <- "Overall"

sample_count <- rbind(sample_count2, sample_count)

sample_count %>%
  gt() %>%
  tab_header(
    title = "Overview of COVID-19 Sequencing Samples")
 

```

***Table 1.1*** Month by Month Sample Counts by Received Source

Total samples processed during each month, for each source/project.

```{r}
month_sample_count <- data_file %>% group_by(received_source, year(PlateDate), month(PlateDate)) %>% summarize(total_samples = length(sample_id))

month_sample_count$ProcessedMonth <- paste0(month_sample_count$`year(PlateDate)`, "-", month_sample_count$`month(PlateDate)`, "-01")

month_sample_count <- dcast(month_sample_count, received_source ~ ProcessedMonth, value.var = c("total_samples"))

month_sample_count[is.na(month_sample_count)] <- 0


###########
# do same thing, but for overall too

ov_sample_count <- data_file %>% group_by(year(PlateDate), month(PlateDate)) %>% summarize(received_source = "Overall", total_samples = length(sample_id))

ov_sample_count$ProcessedMonth <- paste0(ov_sample_count$`year(PlateDate)`, "-", ov_sample_count$`month(PlateDate)`, "-01")

ov_sample_count <- dcast(ov_sample_count, received_source ~ ProcessedMonth, value.var = c("total_samples"))

ov_sample_count[is.na(ov_sample_count)] <- 0

all_month_sample_count <- rbind(ov_sample_count, month_sample_count)

colnames(all_month_sample_count)[1] <- " "

all_month_sample_count %>%
  gt() %>%
  tab_header(
    title = "COVID-19 Sequencing Samples Processed by Month & Received Source")
 

```

***Table 2.*** Status of Pangolin Calls

```{r}
insuf_pang <- data_file %>% group_by(received_source, pangolin_status) %>% summarize(count = sum(counter, na.rm = TRUE))

insuf_pang <- dcast(insuf_pang, received_source ~ pangolin_status, value.var = c("count"))
insuf_pang[is.na(insuf_pang)] <- 0

insuf_pang %>%
  gt() %>%
  tab_header(
    title = "Summary of Pangolin Call Status Messages by Received Source")
```

***Table 3.*** NextClade Completeness by Received Source

```{r}
nc_complete <- data_file %>% group_by(received_source) %>% 
  summarize(minComp = round(min(as.numeric(nextclade_completeness), na.rm = TRUE), 2), 
            tfComp = round(quantile(as.numeric(nextclade_completeness), c(.25), na.rm = TRUE), 2),  
            medComp = round(median(as.numeric(nextclade_completeness), na.rm = TRUE), 2),
            meanComp = round(mean(as.numeric(nextclade_completeness), na.rm = TRUE), 2), 
            sfComp = round(quantile(as.numeric(nextclade_completeness), c(.75), na.rm = TRUE), 2), 
            maxComp = round(max(as.numeric(nextclade_completeness), na.rm = TRUE), 2))

colnames(nc_complete) <- c("Source", "Minimum", "25th Percentile", "Median", "Mean", "75th Percentile", "Maximum")

nc_complete2 <- data_file %>% group_by(counter) %>% 
  summarize(minComp = round(min(as.numeric(nextclade_completeness), na.rm = TRUE), 2), 
            tfComp = round(quantile(as.numeric(nextclade_completeness), c(.25), na.rm = TRUE), 2),  
            medComp = round(median(as.numeric(nextclade_completeness), na.rm = TRUE), 2),
            meanComp = round(mean(as.numeric(nextclade_completeness), na.rm = TRUE), 2), 
            sfComp = round(quantile(as.numeric(nextclade_completeness), c(.75), na.rm = TRUE), 2), 
            maxComp = round(max(as.numeric(nextclade_completeness), na.rm = TRUE), 2))

colnames(nc_complete2) <- c("Source", "Minimum", "25th Percentile", "Median", "Mean", "75th Percentile", "Maximum")

nc_complete2$Source <- "Overall"

nc_complete <- rbind(nc_complete2, nc_complete)

nc_complete %>%
  gt() %>%
  tab_header(
    title = "Summary of Nextclade Completeness (%) by Received Source")
```

***Table 3.1*** Nextclade Completeness by Plate Run - Over 80% Complete

```{r}
plate_complete <- data_file %>% mutate(over80 = case_when(nextclade_completeness >= 80 ~ 1, 
                                                          T ~ 0), 
                                       over90 = case_when(nextclade_completeness >= 90 ~ 1, 
                                                          T ~ 0)) %>% group_by(PlateName, PlateDate, received_source) %>% summarize(over80 = sum(over80, na.rm = TRUE), over90 = sum(over90, na.rm = TRUE), total = sum(counter, na.rm = TRUE))

plate_complete$over80perc <- (plate_complete$over80 / plate_complete$total) * 100
plate_complete$over90perc <- (plate_complete$over90 / plate_complete$total) * 100

ggplot(plate_complete, aes(x = PlateDate, y = over80perc, color = received_source)) + 
  geom_point() + 
  theme_bw() + 
  scale_color_manual(values = c("gray50", "dodgerblue", "dodgerblue", "coral", "seagreen", "goldenrod", "forestgreen", "forestgreen", "goldenrod", "goldenrod", "darkorchid4", "darkorchid4", "darkorchid4")) + 
  labs(title = "Percent of Samples with Genome Completeness Over 80%", 
       x = "Date", 
       y = "Percent (%)", 
       color = "Source", 
       caption = "Each point is a plate/source combination") 

```

***Table 3.2*** Nextclade Completeness by Plate Run - Over 90% Complete

```{r}

ggplot(plate_complete, aes(x = PlateDate, y = over90perc, color = received_source)) + 
  geom_point() + 
  theme_bw() + 
  scale_color_manual(values = c("gray50", "dodgerblue", "dodgerblue", "coral", "seagreen", "goldenrod", "forestgreen", "forestgreen", "goldenrod", "goldenrod", "darkorchid4", "darkorchid4", "darkorchid4")) + 
  labs(title = "Percent of Samples with Genome Completeness Over 90%", 
       x = "Date", 
       y = "Percent (%)", 
       color = "Source", 
       caption = "Each point is a plate/source combination")

```

***Table 4.*** GISAID Information by Received Source

For samples that have a NextClade Completeness value > 90

```{r}
gisaid_info <- filter(data_file, nextclade_completeness > 90) %>% mutate(gs_there = ifelse(gisaid_strain == "", "Blank", "Complete")) %>% group_by(received_source, gs_there) %>% summarize(count = sum(counter, na.rm = TRUE))

gisaid_info <- dcast(gisaid_info, received_source ~ gs_there, value.var = c("count"))

gisaid_info[is.na(gisaid_info)] <- 0

gisaid_info$perc <- round((gisaid_info$Blank / (gisaid_info$Blank + gisaid_info$Complete))*100, 2)

colnames(gisaid_info) <- c("Source", "No GISAID Info", "GISAID Info", "Percent Missing GISAID")

gisaid_info %>%
  gt() %>%
  tab_header(
    title = "Summary of GISAID Completeness by Received Source", 
    subtitle = "Of Samples with NextClade Completeness > 90%")
```

***Table 5. WHO VOIs and VOCs***

(As of WHO site on 7/14/2021)

```{r}
who_voic <- data_file %>% group_by(received_source, voi_voc) %>% summarize(count = sum(counter, na.rm = TRUE))

#table(data_file$voi_voc, data_file$voi_voc_nextclade)

who_voic <- dcast(who_voic, received_source ~ voi_voc, value.var = c("count"))
who_voic[is.na(who_voic)] <- 0

who_voic$Total <- who_voic$`Not VOI/VOC` + who_voic$`VOI/VOC`

who_voic$PercentVOIVOC <- paste0(round((who_voic$`VOI/VOC` / who_voic$Total)*100, 1), "%")

who_voic %>%
  gt() %>%
  tab_header(
    title = "WHO VOIs and VOCs")
```

#### Complete Data Review - Charts

***Chart 1.*** Platforms Used (by Plate Run Date)

```{r}
platforms <- data_file %>% group_by(PlateDate, PlatePlatform) %>% summarize(count = sum(counter, na.rm = TRUE))

ill_sum <- sum(filter(platforms, PlatePlatform == "Illumina")$count, na.rm = TRUE)
nan_sum <- sum(filter(platforms, PlatePlatform == "Nanopore")$count, na.rm = TRUE)

ggplot(platforms, aes(x = PlateDate, y = count, color = PlatePlatform)) + 
  geom_point(alpha = 0.5) + 
  geom_line() + 
  theme_bw() + 
  labs(title = "Samples Run on Various Platforms", 
       x = "Plate Run Date", 
       y = "Count of Samples Run", 
       color = "Platforms", 
       caption = paste0("Total Samples Run with Illumina = ", ill_sum, "; Total Samples Run with Nanopore = ", nan_sum, ";"))


```

***Chart 1.1*** Platforms Used by Received Source (by Plate Run Date)

```{r}
platforms2 <- data_file %>% group_by(PlateDate, PlatePlatform, received_source) %>% summarize(count = sum(counter, na.rm = TRUE))

ggplot(platforms2, aes(x = PlateDate, y = count, color = received_source)) + 
  geom_point(alpha = 0.7, aes(shape = platforms2$PlatePlatform)) + 
  #geom_line() + 
  theme_bw() + 
  labs(title = "Samples Run on Various Platforms", 
       subtitle = "by Received Source",
       x = "Plate Run Date", 
       y = "Count of Samples Run", 
       color = "Source", 
       shape = "Platform")


```


***Chart 2.*** Pangolin Lineage Calls Over Time (Sample Collection)

```{r}
pangolin_calls <- data_file %>% group_by(coll_date, pangolin_lineage, voi_voc_pangolin) %>% summarize(count = sum(counter, na.rm = TRUE), ind_count = length(unique(sample_id)))

pangolin_calls <- filter(pangolin_calls, coll_date < Sys.Date())

ggplot(pangolin_calls, aes(x = coll_date, y = ind_count, fill = voi_voc_pangolin)) + 
  geom_bar(stat = "identity") + 
  theme_bw() +
  scale_fill_manual(values = c("dodgerblue", "dodgerblue", "coral2", "darkcyan", "darkcyan", 
                               "darkcyan", "deepskyblue", "deepskyblue4", "darkolivegreen2", 
                               "darkolivegreen3", "forestgreen", "forestgreen", "darkorchid", 
                               "gray75", "mediumpurple1", "mediumpurple1", "mediumpurple1", 
                               "lightskyblue", "plum4")) + 
  #theme(legend.position = "none") +
  labs(title = "Pangolin Lineage Calls Over Time", 
       x = "Sample Collection Date", 
       y = "Sample Count", 
       fill = "WHO VOIs and VOCs")

```


***Chart 2.1*** Pangolin Lineage Calls by Received Source + Date of Sample Collection

```{r, fig.height = 20, fig.width = 8}
pangolin_calls2 <- data_file %>% group_by(coll_date, pangolin_lineage, received_source, voi_voc_pangolin) %>% summarize(count = sum(counter, na.rm = TRUE), ind_count = length(unique(sample_id)))

pangolin_calls2 <- filter(pangolin_calls2, coll_date < Sys.Date())

ggplot(pangolin_calls2, aes(x = coll_date, y = ind_count, fill = voi_voc_pangolin)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + 
  scale_fill_manual(values = c("dodgerblue", "dodgerblue", "coral2", "darkcyan", "darkcyan", 
                               "darkcyan", "deepskyblue", "deepskyblue4", "darkolivegreen2", 
                               "darkolivegreen3", "forestgreen", "forestgreen", "darkorchid", 
                               "gray75", "mediumpurple1", "mediumpurple1", "mediumpurple1", 
                               "lightskyblue", "plum4")) + 
  #theme(legend.position = "none") +
  labs(title = "Pangolin Lineage Calls Over Time",
       subtitle = "by Received Source",
       x = "Sample Collection Date", 
       y = "Sample Count", 
       fill = "WHO VOIs and VOCs") + 
  facet_wrap(.~received_source, ncol = 2)

```


***Chart 3.*** NextClade Clade Calls by Date of Sample Collection

```{r}
nextclade_calls <- data_file %>% group_by(coll_date, nextclade_clade, voi_voc_nextclade) %>% summarize(ind_count = length(unique(sample_id)))

nextclade_calls <- filter(nextclade_calls, coll_date < Sys.Date())

ggplot(nextclade_calls, aes(x = coll_date, y = ind_count, fill = voi_voc_nextclade)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + 
  #theme(legend.position = "none") +
  scale_fill_manual(values = c("dodgerblue", "mediumpurple", "goldenrod", "orange3", "violetred", "tomato3", "skyblue1", "slateblue1", "palegreen3", "darkorchid4", "lightblue1", "gray75", "olivedrab3", "lightgoldenrod1", "lightpink")) + 
  labs(title = "NextClade Clade Calls Over Time", 
       x = "Sample Collection Date", 
       y = "Sample Count", 
       fill = "WHO VOIs and VOCs")

```

***Chart 3.1*** NextClade Clade Calls by Received Source + Date of Sample Collection

```{r, fig.height = 20, fig.width = 8}
nextclade_calls2 <- data_file %>% group_by(received_source, coll_date, nextclade_clade, voi_voc_nextclade) %>% summarize(ind_count = length(unique(sample_id)))

nextclade_calls2 <- filter(nextclade_calls2, coll_date < Sys.Date())

ggplot(nextclade_calls2, aes(x = coll_date, y = ind_count, fill = voi_voc_nextclade)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + 
  #theme(legend.position = "none") +
  scale_fill_manual(values = c("dodgerblue", "mediumpurple", "goldenrod", "orange3", "violetred", "tomato3", "skyblue1", "slateblue1", "palegreen3", "darkorchid4", "lightblue1", "gray75", "olivedrab3", "lightgoldenrod1", "lightpink")) + 
  labs(title = "NextClade Clade Calls Over Time", 
       subtitle = "by Received Source",
       x = "Sample Collection Date", 
       y = "Sample Count", 
       fill = "WHO VOIs and VOCs") + 
  facet_wrap(.~received_source, ncol = 2)

```


***Chart 4.*** GISAID Strains Over Time (Sample Collection)

```{r}

gisaid_strains <- data_file %>% group_by(coll_date, gisaid_strain) %>% summarize(ind_count = length(unique(sample_id)))

gisaid_strains <- filter(gisaid_strains, coll_date < Sys.Date())

ggplot(gisaid_strains, aes(x = coll_date, y = ind_count, fill = gisaid_strain)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + 
  theme(legend.position = "none") +
  labs(title = "GISAID Strain Calls Over Time", 
       x = "Sample Collection Date", 
       y = "Sample Count")

```



***Chart 4.1*** GISAID Strains by Received Source + Week of Sample Collection

```{r, fig.height = 20, fig.width = 8}

gisaid_strains2 <- data_file %>% group_by(received_source, coll_date, gisaid_strain) %>% summarize(ind_count = length(unique(sample_id)))

gisaid_strains2 <- filter(gisaid_strains2, coll_date < Sys.Date())

ggplot(gisaid_strains2, aes(x = coll_date, y = ind_count, fill = gisaid_strain)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + 
  theme(legend.position = "none") +
  labs(title = "GISAID Strain Calls Over Time", 
       subtitle = "by Received Source",
       x = "Sample Collection Date", 
       y = "Sample Count") + 
  facet_wrap(.~received_source, ncol = 2)

```

***Chart 5.*** Common Pangolin Calls by Week of Sample Collection

```{r}
common_pangolin_calls <- data_file %>% group_by(received_source, coll_date, pangolin_lineage) %>% summarize(count = sum(counter, na.rm = TRUE), ind_count = length(unique(sample_id)))

common_pangolin_calls <- filter(common_pangolin_calls, coll_date < Sys.Date())

common_pangolin_calls$week <- paste0(year(common_pangolin_calls$coll_date), "-", ifelse(nchar(epiweek(common_pangolin_calls$coll_date)) == 1, paste0("0", as.character(epiweek(common_pangolin_calls$coll_date))), epiweek(common_pangolin_calls$coll_date)))

# week is counting last week of 2020 as week "53"
common_pangolin_calls$week <- ifelse(common_pangolin_calls$week == "2021-53" & year(common_pangolin_calls$coll_date) == 2021, "2021-01", common_pangolin_calls$week)

common_pangolin_calls <- common_pangolin_calls %>% group_by(week, received_source) %>% top_n(1, ind_count)

ggplot(common_pangolin_calls, aes(x = week, y = pangolin_lineage, color = received_source)) +   
  geom_point(size = common_pangolin_calls$ind_count, alpha = 0.5) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Most Common Pangolin Lineages by Week", 
       x = "Collection Date", 
       y = "Lineage", 
       color = "Source", 
       caption = "Points are sized based on the number of samples")
```

***Chart 5.*** Common NextClade Calls by Week of Sample Collection

```{r}
common_nextclade_calls <- data_file %>% group_by(received_source, coll_date, nextclade_clade) %>% summarize(count = sum(counter, na.rm = TRUE), ind_count = length(unique(sample_id)))

common_nextclade_calls <- filter(common_nextclade_calls, coll_date < Sys.Date())

common_nextclade_calls$week <- paste0(year(common_nextclade_calls$coll_date), "-", ifelse(nchar(epiweek(common_nextclade_calls$coll_date)) == 1, paste0("0", as.character(epiweek(common_nextclade_calls$coll_date))), epiweek(common_nextclade_calls$coll_date)))

# week is counting last week of 2020 as week "53"
common_nextclade_calls$week <- ifelse(common_nextclade_calls$week == "2021-53" & year(common_nextclade_calls$coll_date) == 2021, "2021-01", common_nextclade_calls$week)


common_nextclade_calls <- common_nextclade_calls %>% group_by(week, received_source) %>% top_n(1, ind_count)

ggplot(common_nextclade_calls, aes(x = week, y = nextclade_clade, color = received_source)) +   
  geom_point(size = common_nextclade_calls$ind_count, alpha = 0.5) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Most Common NextClade Clades by Week", 
       x = "Collection Date", 
       y = "Lineage", 
       color = "Source", 
       caption = "Points are sized based on the number of samples")
```

***Chart 6.*** NextClade Completeness Over Time by Received Source + Week of Plate Run

```{r, fig.height = 15, fig.width = 8}
data_file$week <- paste0(year(data_file$PlateDate), "-", ifelse(nchar(epiweek(data_file$PlateDate)) == 1, paste0("0", as.character(epiweek(data_file$PlateDate))), epiweek(data_file$PlateDate)))

# week is counting last week of 2020 as week "53"
data_file$week <- ifelse(data_file$week == "2021-53" & year(data_file$coll_date) == 2021, "2021-01", data_file$week)

ggplot(data_file, aes(x = week, y = as.numeric(nextclade_completeness))) + 
  geom_boxplot() + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "NextClade Completeness by Week of Plate Run", 
       x = "Week", 
       y = "Completeness (%)") + 
  facet_wrap(.~received_source, ncol = 2)

```


***Chart 6.1*** NextClade Completeness Over Time by Received Source + Week of Sample Collection

```{r, fig.height = 15, fig.width = 8}
data_file$week2 <- paste0(year(data_file$coll_date), "-", ifelse(nchar(epiweek(data_file$coll_date)) == 1, paste0("0", as.character(epiweek(data_file$coll_date))), epiweek(data_file$coll_date)))

# week is counting last week of 2020 as week "53"
data_file$week2 <- ifelse(data_file$week2 == "2021-53" & year(data_file$coll_date) == 2021, "2021-01", data_file$week2)


ggplot(data_file, aes(x = week2, y = as.numeric(nextclade_completeness))) + 
  geom_boxplot() + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "NextClade Completeness by Week of Sample Collection", 
       x = "Week", 
       y = "Completeness (%)") + 
  facet_wrap(.~received_source, ncol = 2)

```


***Chart 7.*** Percent of Calls = B117 per Week (by collection date)

based on Pangolin Calls

```{r}
pangolin_b117 <- data_file %>% group_by(week2, coll_date, pangolin_lineage) %>% summarize(count = sum(counter, na.rm = TRUE), ind_count = length(unique(sample_id)))

pangolin_b117$b117yn <- ifelse(pangolin_b117$pangolin_lineage == "B.1.1.7", "B117", "Other")

pangolin_b117 <- pangolin_b117 %>% group_by(week2, b117yn) %>% summarize(count_all = sum(count, na.rm = TRUE))

pangolin_b117 <- dcast(pangolin_b117, week2 ~ b117yn, value.var = c("count_all"))

pangolin_b117[is.na(pangolin_b117)] <- 0

pangolin_b117$percent <- (pangolin_b117$B117 / (pangolin_b117$B117 + pangolin_b117$Other))*100
pangolin_b117$total <- (pangolin_b117$B117 + pangolin_b117$Other)

ggplot(pangolin_b117, aes(x = week2, y = percent)) + 
  geom_point() + 
  geom_line() + 
  geom_text(aes(x = week2, y = percent, label = total), nudge_y = 3, size = 3) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Percent of Samples Collected Each Week Identified as B.1.1.7", 
       x = "Week", 
       y = "Percent (%)", 
       caption = "Pangolin used to determine strain; Number label = Count of samples collected each week")
  
```

Should we do anything with: nextclade_qcOverallScore, nextclade_qcOverallStatus, nextclade_totalMutations, nextclade_totalNonACGTNs?

