---
title: "Sequence Report"
author: "Julie Gilbert"
date: "6/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

# Report Overview

This report uses ```/SequenceSampleMetadata/FinalSummary/secret/full_compiled_data.csv``` as its data source. The goal of this report is to provide summary statistics on the COVID-19 sequencing status of Dr. Adam Lauring's research lab. 

This report was run on `r Sys.Date()`.

For questions, please contact Julie (Jules) Gilbert [juliegil@umich.edu].

--- 

```{r}
library(tidyverse)
library(reshape2)
library(lubridate)
library(gt)
```

```{r}

data_file <- read.csv("C:/Users/juliegil/Dropbox (University of Michigan)/MED-LauringLab/SequenceSampleMetadata/FinalSummary/secret/full_compiled_data.csv", colClasses = c("character"))

```


## Complete Data Review

```{r}
data_file$coll_date <- as_date(data_file$coll_date)

data_file$PlateDate <- as_date(data_file$PlateDate)

data_file$counter <- 1
```


The full_compiled_data.csv dataset includes `r nrow(data_file)` sample records. This includes samples collected from `r min(data_file$coll_date)` to `r max(data_file$coll_date)`. The most recent samples were processed on `r max(data_file$PlateDate)`.

***Table 1.*** Sample Count from Received Sources

```{r}
sample_count <- data_file %>% group_by(received_source) %>% summarize(unique_samples = length(unique(sample_id)), total_samples = length(sample_id), unique_subjects = length(unique(subject_id)), min_collection = min(coll_date), max_collection = max(coll_date), min_plate_run = min(PlateDate), max_plate_run = max(PlateDate))

colnames(sample_count) <- c("Source", "Count of Unique Samples", "Count of Total Samples", "Count of Unique Subjects", "Min. Collection Date", "Max. Collection Date", "Min. Plate Run Date", "Max. Plate Run Date")

sample_count2 <- data_file %>% group_by(counter) %>% summarize(unique_samples = length(unique(sample_id)), total_samples = length(sample_id), unique_subjects = length(unique(subject_id)), min_collection = min(coll_date), max_collection = max(coll_date), min_plate_run = min(PlateDate), max_plate_run = max(PlateDate))

colnames(sample_count2) <- c("Source", "Count of Unique Samples", "Count of Total Samples", "Count of Unique Subjects", "Min. Collection Date", "Max. Collection Date", "Min. Plate Run Date", "Max. Plate Run Date")

sample_count2$Source <- "Overall"

sample_count <- rbind(sample_count2, sample_count)

sample_count %>%
  gt() %>%
  tab_header(
    title = "Overview of COVID-19 Sequencing Samples")
 

```


***Table 2.*** Status of Pangolin Calls

```{r}
insuf_pang <- data_file %>% group_by(received_source, pangolin_status) %>% summarize(count = sum(counter, na.rm = TRUE))

insuf_pang <- dcast(insuf_pang, received_source ~ pangolin_status, value.var = c("count"))
insuf_pang[is.na(insuf_pang)] <- 0

insuf_pang %>%
  gt() %>%
  tab_header(
    title = "Summary of Pangolin Call Status Messages by Received Source")
```

***Table 3.*** NextClade Completeness by Received Source

```{r}
nc_complete <- data_file %>% group_by(received_source) %>% 
  summarize(minComp = round(min(as.numeric(nextclade_completeness), na.rm = TRUE), 2), 
            tfComp = round(quantile(as.numeric(nextclade_completeness), c(.25), na.rm = TRUE), 2),  
            medComp = round(median(as.numeric(nextclade_completeness), na.rm = TRUE), 2),
            meanComp = round(mean(as.numeric(nextclade_completeness), na.rm = TRUE), 2), 
            sfComp = round(quantile(as.numeric(nextclade_completeness), c(.75), na.rm = TRUE), 2), 
            maxComp = round(max(as.numeric(nextclade_completeness), na.rm = TRUE), 2))

colnames(nc_complete) <- c("Source", "Minimum", "25th Percentile", "Median", "Mean", "75th Percentile", "Maximum")

nc_complete2 <- data_file %>% group_by(counter) %>% 
  summarize(minComp = round(min(as.numeric(nextclade_completeness), na.rm = TRUE), 2), 
            tfComp = round(quantile(as.numeric(nextclade_completeness), c(.25), na.rm = TRUE), 2),  
            medComp = round(median(as.numeric(nextclade_completeness), na.rm = TRUE), 2),
            meanComp = round(mean(as.numeric(nextclade_completeness), na.rm = TRUE), 2), 
            sfComp = round(quantile(as.numeric(nextclade_completeness), c(.75), na.rm = TRUE), 2), 
            maxComp = round(max(as.numeric(nextclade_completeness), na.rm = TRUE), 2))

colnames(nc_complete2) <- c("Source", "Minimum", "25th Percentile", "Median", "Mean", "75th Percentile", "Maximum")

nc_complete2$Source <- "Overall"

nc_complete <- rbind(nc_complete2, nc_complete)

nc_complete %>%
  gt() %>%
  tab_header(
    title = "Summary of Nextclade Completeness (%) by Received Source")
```


***Table 4.*** GISAID Information by Received Source

For samples that have a NextClade Completeness value > 90

```{r}
gisaid_info <- filter(data_file, nextclade_completeness > 90) %>% mutate(gs_there = ifelse(gisaid_strain == "", "Blank", "Complete")) %>% group_by(received_source, gs_there) %>% summarize(count = sum(counter, na.rm = TRUE))

gisaid_info <- dcast(gisaid_info, received_source ~ gs_there, value.var = c("count"))

gisaid_info[is.na(gisaid_info)] <- 0

gisaid_info$perc <- round((gisaid_info$Blank / (gisaid_info$Blank + gisaid_info$Complete))*100, 2)

colnames(gisaid_info) <- c("Source", "No GISAID Info", "GISAID Info", "Percent Missing GISAID")

gisaid_info %>%
  gt() %>%
  tab_header(
    title = "Summary of GISAID Completeness by Received Source", 
    subtitle = "Of Samples with NextClade Completeness > 90%")
```

#### Complete Data Review - Charts

***Chart 1.*** Platforms Used (by Plate Run Date)

```{r}
platforms <- data_file %>% group_by(PlateDate, PlatePlatform) %>% summarize(count = sum(counter, na.rm = TRUE))

ill_sum <- sum(filter(platforms, PlatePlatform == "Illumina")$count, na.rm = TRUE)
nan_sum <- sum(filter(platforms, PlatePlatform == "Nanopore")$count, na.rm = TRUE)

ggplot(platforms, aes(x = PlateDate, y = count, color = PlatePlatform)) + 
  geom_point(alpha = 0.5) + 
  geom_line() + 
  theme_bw() + 
  labs(title = "Samples Run on Various Platforms", 
       x = "Plate Run Date", 
       y = "Count of Samples Run", 
       color = "Platforms", 
       caption = paste0("Total Samples Run with Illumina = ", ill_sum, "; Total Samples Run with Nanopore = ", nan_sum, ";"))


```

***Chart 1.1*** Platforms Used by Received Source (by Plate Run Date)

```{r}
platforms2 <- data_file %>% group_by(PlateDate, PlatePlatform, received_source) %>% summarize(count = sum(counter, na.rm = TRUE))

ggplot(platforms2, aes(x = PlateDate, y = count, color = received_source)) + 
  geom_point(alpha = 0.7, aes(shape = platforms2$PlatePlatform)) + 
  #geom_line() + 
  theme_bw() + 
  labs(title = "Samples Run on Various Platforms", 
       subtitle = "by Received Source",
       x = "Plate Run Date", 
       y = "Count of Samples Run", 
       color = "Source", 
       shape = "Platform")


```


***Chart 2.*** Pangolin Lineage Calls Over Time (Sample Collection)

```{r}
pangolin_calls <- data_file %>% group_by(coll_date, pangolin_lineage) %>% summarize(count = sum(counter, na.rm = TRUE), ind_count = length(unique(sample_id)))

pangolin_calls <- filter(pangolin_calls, coll_date < Sys.Date())

ggplot(pangolin_calls, aes(x = coll_date, y = ind_count, fill = pangolin_lineage)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + 
  theme(legend.position = "none") +
  labs(title = "Pangolin Lineage Calls Over Time", 
       x = "Sample Collection Date", 
       y = "Sample Count")

```


***Chart 2.1*** Pangolin Lineage Calls by Received Source + Date of Sample Collection

```{r}
pangolin_calls2 <- data_file %>% group_by(coll_date, pangolin_lineage, received_source) %>% summarize(count = sum(counter, na.rm = TRUE), ind_count = length(unique(sample_id)))

pangolin_calls2 <- filter(pangolin_calls2, coll_date < Sys.Date())

ggplot(pangolin_calls2, aes(x = coll_date, y = ind_count, fill = pangolin_lineage)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + 
  theme(legend.position = "none") +
  labs(title = "Pangolin Lineage Calls Over Time",
       subtitle = "by Received Source",
       x = "Sample Collection Date", 
       y = "Sample Count") + 
  facet_wrap(.~received_source)

```


***Chart 3.*** NextClade Clade Calls by Date of Sample Collection

```{r}
nextclade_calls <- data_file %>% group_by(coll_date, nextclade_clade) %>% summarize(ind_count = length(unique(sample_id)))

nextclade_calls <- filter(nextclade_calls, coll_date < Sys.Date())

ggplot(nextclade_calls, aes(x = coll_date, y = ind_count, fill = nextclade_clade)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + 
  theme(legend.position = "none") +
  labs(title = "NextClade Clade Calls Over Time", 
       x = "Sample Collection Date", 
       y = "Sample Count")

```

***Chart 3.1*** NextClade Clade Calls by Received Source + Date of Sample Collection

```{r}
nextclade_calls2 <- data_file %>% group_by(received_source, coll_date, nextclade_clade) %>% summarize(ind_count = length(unique(sample_id)))

nextclade_calls2 <- filter(nextclade_calls2, coll_date < Sys.Date())

ggplot(nextclade_calls2, aes(x = coll_date, y = ind_count, fill = nextclade_clade)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + 
  theme(legend.position = "none") +
  labs(title = "NextClade Clade Calls Over Time", 
       subtitle = "by Received Source",
       x = "Sample Collection Date", 
       y = "Sample Count") + 
  facet_wrap(.~received_source)

```


***Chart 4.*** GISAID Strains Over Time (Sample Collection)

```{r}

gisaid_strains <- data_file %>% group_by(coll_date, gisaid_strain) %>% summarize(ind_count = length(unique(sample_id)))

gisaid_strains <- filter(gisaid_strains, coll_date < Sys.Date())

ggplot(gisaid_strains, aes(x = coll_date, y = ind_count, fill = gisaid_strain)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + 
  theme(legend.position = "none") +
  labs(title = "GISAID Strain Calls Over Time", 
       x = "Sample Collection Date", 
       y = "Sample Count")

```



***Chart 4.1*** GISAID Strains by Received Source + Week of Sample Collection

```{r}

gisaid_strains2 <- data_file %>% group_by(received_source, coll_date, gisaid_strain) %>% summarize(ind_count = length(unique(sample_id)))

gisaid_strains2 <- filter(gisaid_strains2, coll_date < Sys.Date())

ggplot(gisaid_strains2, aes(x = coll_date, y = ind_count, fill = gisaid_strain)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + 
  theme(legend.position = "none") +
  labs(title = "GISAID Strain Calls Over Time", 
       subtitle = "by Received Source",
       x = "Sample Collection Date", 
       y = "Sample Count") + 
  facet_wrap(.~received_source)

```

***Chart 5.*** Common Pangolin Calls by Week of Sample Collection

```{r}
common_pangolin_calls <- data_file %>% group_by(received_source, coll_date, pangolin_lineage) %>% summarize(count = sum(counter, na.rm = TRUE), ind_count = length(unique(sample_id)))

common_pangolin_calls <- filter(common_pangolin_calls, coll_date < Sys.Date())

common_pangolin_calls$week <- paste0(year(common_pangolin_calls$coll_date), "-", ifelse(nchar(epiweek(common_pangolin_calls$coll_date)) == 1, paste0("0", as.character(epiweek(common_pangolin_calls$coll_date))), epiweek(common_pangolin_calls$coll_date)))

common_pangolin_calls <- common_pangolin_calls %>% group_by(week, received_source) %>% top_n(1, ind_count)

ggplot(common_pangolin_calls, aes(x = week, y = pangolin_lineage, color = received_source)) +   
  geom_point(size = common_pangolin_calls$ind_count, alpha = 0.5) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Most Common Pangolin Lineages by Week", 
       x = "Collection Date", 
       y = "Lineage", 
       color = "Source", 
       caption = "Points are sized based on the number of samples")
```

***Chart 5.*** Common NextClade Calls by Week of Sample Collection

```{r}
common_nextclade_calls <- data_file %>% group_by(received_source, coll_date, nextclade_clade) %>% summarize(count = sum(counter, na.rm = TRUE), ind_count = length(unique(sample_id)))

common_nextclade_calls <- filter(common_nextclade_calls, coll_date < Sys.Date())

common_nextclade_calls$week <- paste0(year(common_nextclade_calls$coll_date), "-", ifelse(nchar(epiweek(common_nextclade_calls$coll_date)) == 1, paste0("0", as.character(epiweek(common_nextclade_calls$coll_date))), epiweek(common_nextclade_calls$coll_date)))

common_nextclade_calls <- common_nextclade_calls %>% group_by(week, received_source) %>% top_n(1, ind_count)

ggplot(common_nextclade_calls, aes(x = week, y = nextclade_clade, color = received_source)) +   
  geom_point(size = common_nextclade_calls$ind_count, alpha = 0.5) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Most Common NextClade Clades by Week", 
       x = "Collection Date", 
       y = "Lineage", 
       color = "Source", 
       caption = "Points are sized based on the number of samples")
```

***Chart 6.*** NextClade Completeness Over Time by Received Source + Week of Plate Run

```{r}
data_file$week <- paste0(year(data_file$PlateDate), "-", ifelse(nchar(epiweek(data_file$PlateDate)) == 1, paste0("0", as.character(epiweek(data_file$PlateDate))), epiweek(data_file$PlateDate)))

ggplot(data_file, aes(x = week, y = as.numeric(nextclade_completeness))) + 
  geom_boxplot() + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "NextClade Completeness by Week of Plate Run", 
       x = "Week", 
       y = "Completeness (%)") + 
  facet_wrap(.~received_source, ncol = 2)

```


***Chart 6.1*** NextClade Completeness Over Time by Received Source + Week of Sample Collection

```{r}
data_file$week2 <- paste0(year(data_file$coll_date), "-", ifelse(nchar(epiweek(data_file$coll_date)) == 1, paste0("0", as.character(epiweek(data_file$coll_date))), epiweek(data_file$coll_date)))

ggplot(data_file, aes(x = week2, y = as.numeric(nextclade_completeness))) + 
  geom_boxplot() + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title = "NextClade Completeness by Week of Sample Collection", 
       x = "Week", 
       y = "Completeness (%)") + 
  facet_wrap(.~received_source, ncol = 2)

```



Should we do anything with: nextclade_qcOverallScore, nextclade_qcOverallStatus, nextclade_totalMutations, nextclade_totalNonACGTNs?

