---
title: "Sequence Report"
author: "Julie Gilbert"
date: "6/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Report Overview

This report uses ```/SequenceSampleMetadata/FinalSummary/secret/full_compiled_data.csv``` as its data source. The goal of this report is to provide summary statistics on the COVID-19 sequencing status of Dr. Adam Lauring's research lab. 

This report was run on `r Sys.Date()`.

For questions, please contact Julie (Jules) Gilbert [juliegil@umich.edu].

--- 

```{r}
library(tidyverse)
library(reshape2)
library(lubridate)
library(gt)
```

```{r}

data_file <- read.csv("C:/Users/juliegil/Dropbox (University of Michigan)/MED-LauringLab/SequenceSampleMetadata/FinalSummary/secret/full_compiled_data.csv", colClasses = c("character"))

```


## Complete Data Review

```{r}
data_file$coll_date <- as_date(data_file$coll_date)

data_file$PlateDate <- as_date(data_file$PlateDate)

data_file$counter <- 1
```


The full_compiled_data.csv dataset includes `r nrow(data_file)` sample records. This includes samples collected from `r min(data_file$coll_date)` to `r max(data_file$coll_date)`. The most recent samples were processed on `r max(data_file$PlateDate)`.

***Table 1.*** Sample Count from Received Sources

```{r}
sample_count <- data_file %>% group_by(received_source) %>% summarize(unique_samples = length(unique(sample_id)), total_samples = length(sample_id), unique_subjects = length(unique(subject_id)), min_collection = min(coll_date), max_collection = max(coll_date), min_plate_run = min(PlateDate), max_plate_run = max(PlateDate))

colnames(sample_count) <- c("Source", "Count of Unique Samples", "Count of Total Samples", "Count of Unique Subjects", "Min. Collection Date", "Max. Collection Date", "Min. Plate Run Date", "Max. Plate Run Date")

sample_count %>%
  gt() %>%
  tab_header(
    title = "Overview of COVID-19 Sequencing Samples")
 

```

Add: 

Table of Count - insufficient pangolin calls (pangolin probability/status)
Table of most common pangolin call per week
Table of Nextclade 5 number summary on completeness by received source + overall
Table of most common next clade clade call per week


#### Complete Data Review - Charts

***Chart 1.*** Platforms Used

```{r}
platforms <- data_file %>% group_by(PlateDate, PlatePlatform) %>% summarize(count = sum(counter, na.rm = TRUE))

ill_sum <- sum(filter(platforms, PlatePlatform == "Illumina")$count, na.rm = TRUE)
nan_sum <- sum(filter(platforms, PlatePlatform == "Nanopore")$count, na.rm = TRUE)

ggplot(platforms, aes(x = PlateDate, y = count, color = PlatePlatform)) + 
  geom_point(alpha = 0.5) + 
  geom_line() + 
  theme_bw() + 
  labs(title = "Samples Run on Various Platforms", 
       x = "Plate Run Date", 
       y = "Count of Samples Run", 
       color = "Platforms", 
       caption = paste0("Total Samples Run with Illumina = ", ill_sum, "; Total Samples Run with Nanopore = ", nan_sum, ";"))


```

***Chart 2.*** Pangolin Lineage Calls Over Time

```{r}
pangolin_calls <- data_file %>% group_by(coll_date, pangolin_lineage) %>% summarize(count = sum(counter, na.rm = TRUE), ind_count = length(unique(sample_id)))

pangolin_calls <- filter(pangolin_calls, coll_date < Sys.Date())

ggplot(pangolin_calls, aes(x = coll_date, y = ind_count, fill = pangolin_lineage)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + 
  theme(legend.position = "none") +
  labs(title = "Pangolin Lineage Calls Over Time", 
       x = "Sample Collection Date", 
       y = "Sample Count")

```

***Chart 3.*** NextClade Clade Calls Over Time

```{r}
nextclade_calls <- data_file %>% group_by(coll_date, nextclade_clade) %>% summarize(ind_count = length(unique(sample_id)))

nextclade_calls <- filter(nextclade_calls, coll_date < Sys.Date())

ggplot(nextclade_calls, aes(x = coll_date, y = ind_count, fill = nextclade_clade)) + 
  geom_bar(stat = "identity") + 
  theme_bw() + 
  theme(legend.position = "none") +
  labs(title = "NextClade Clade Calls Over Time", 
       x = "Sample Collection Date", 
       y = "Sample Count")

```




Add: 

Chart of Nextclade 5 number summary on completeness by received source + overall (by week)


